---
title: "615 assignment 3 data cleaning"
author: "Si Chen,Grace Xie, Kaiyu yan, Siwei Hu"
date: "September 30, 2018"
output:
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,dev="CairoPNG",fig.align = "center", 
                      fig.width = 5.656, fig.height = 4, global.par = TRUE)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60),  # For code
                      width = 60)  # For output
options(cli.width = 60)  # For tidyverse loading messages
#install.packages("pacman",repos="https://cloud.r-project.org")
pacman::p_load("ggplot2","knitr","arm","foreign","car","Cairo","data.table","dplyr","lattice","RColorBrewer","pheatmap")
par (mar=c(3,3,2,1), mgp=c(2,.7,0), tck=-.01)
```



# Introduction

We collect the data from he Data.gov, named Community Health Status Indicators (CHSI) to Combat Obesity, Heart Disease and Cancer(Chsi_dataset).It is imperative to understand that behavioral factors such as obesity, tobacco use, diet, physical activity, alcohol and drug use, sexual behavior and others substantially contribute to these deaths. After data cleaning, we choose obesity, uninsurence, diabetes, smkoing as our variables. 

```{r }
#import datasets from MEASURESOFBIRTHANDDEATH and RISKFACTORSANDACCESSTOCARE
tableM <- read.csv("MEASURESOFBIRTHANDDEATH.csv")
tableR <- read.csv("RISKFACTORSANDACCESSTOCARE.csv")
tableD <-read.csv("DEMOGRAPHICS.csv")

# select the variable total death in RISKFACTORSANDACCESSTOCARE
death <- select(tableM, Total_Deaths)

# select the variable obesity, smoker, diabetes, uninsured, Dentist_Rate,CHSI_State_Name from RISKFACTORSANDACCESSTOCARE
risk <- select(tableR,Obesity,Smoker,Diabetes,Uninsured,Dentist_Rate,CHSI_State_Name)
#selec the variable population_size and build a new table
population <- select(tableD, Population_Size)

# combine to a new table
tablenew <- cbind.data.frame(death,risk,population)

# remove all useless inputs
tablenew[tablenew < 0] <- NA

# create the final data frame
tablenew1<-tablenew[complete.cases(tablenew),]


```

We visualize the datasets with different units: percentage and counts, have seperately created the ggplot for each variable. More details are listed below.

```{r}
# total population and death in every state

T_population<-tablenew1 %>%
 group_by(CHSI_State_Name) %>%
 summarise(total_p = sum(Population_Size)/10000,total_d=sum(Total_Deaths)/1000,
           death_rate=total_d/total_p) %>%
 arrange(desc(total_p))
T_p<-data.frame(T_population[c(1:10),])
T_p1<-T_p %>% select(CHSI_State_Name,total_d)%>% mutate(type=rep("1",10)) %>% arrange(desc(total_d))
T_p2<-T_p %>% select(CHSI_State_Name,total_p) %>% mutate(type=rep("0",10))
names(T_p1) <- c("State","Number","Type")
names(T_p2) <- c("State","Number","Type")
T_p_f<-merge(T_p1,T_p2,all=T)


ggplot(data=T_p_f, aes(x=State, y=Number, fill=Type)) +
geom_bar(stat="identity",position=position_dodge(),color="black")+
theme_minimal()+
theme(axis.text.x = element_text(angle = 90))+ 
scale_fill_brewer(palette="Paired",labels=c("Total population", "Death population"))+
scale_x_discrete(limits=c("California","Texas","New York", "Florida","Pennsylvania","Illinois","Michigan","Ohio","New Jersey","North Carolina"))+
ggtitle("Top 10 States in toal populaton and death population")


```

We choose Top 10 total population states and filter their relation between population and total deaths to make more sense of the data.   


```{r}
# heat plot 
cols<-brewer.pal(4,"Spectral")
col<-colorRampPalette(cols)
heat_data<-tablenew1 %>% select(Obesity,Smoker,Diabetes,Uninsured,Total_Deaths,Population_Size) %>% 
mutate(uninsured=Uninsured/Population_Size*100,Death=Total_Deaths/Population_Size*100)
heat_data<-heat_data[,-c(4,5,6)]
vc<-cor(heat_data)
levelplot(vc,col.regions=col(100),main="Cor Plot Test",scales=list(x=list(rot=90)))


```

we select all variables which may impact the death rate and scale our uninsured and death data into percent. Then we build a chart about correlation between each two variables. 

```{r }
Obesityp <- tablenew1 %>% 
  select(Obesity,Population_Size,Total_Deaths,CHSI_State_Name) %>% 
  mutate(obesity_p = Obesity/100*Population_Size)

table.state1 <- Obesityp %>% group_by(CHSI_State_Name) %>% summarise(Total_Deaths = sum(Total_Deaths),obesity_p=sum(obesity_p))
#tablenew2 <- cbind.data.frame(Obesityp, tablenew1)

ggplot(table.state1) + geom_point(aes(x = obesity_p, y = Total_Deaths, color = obesity_p)) + geom_smooth(aes(x = obesity_p, y = Total_Deaths)) + ggtitle( "SiChen： Obesity vs Death")


ggplot(tablenew1) + geom_point(aes(x = Obesity, y = Total_Deaths, color = Obesity)) + geom_smooth(aes(x = Obesity, y = Total_Deaths)) + ggtitle( "SiChen： Obesity vs Death")



```


```{r}
table.state2 <- tablenew1 %>% group_by(CHSI_State_Name) %>% summarise(totaldeath = sum(Total_Deaths),uninsured=sum(Uninsured))

ggplot(table.state2)+
  ggtitle("siwei: uninsured VS total_death")+
  geom_point(mapping = aes(x = uninsured,y = totaldeath))+
  geom_smooth(mapping = aes(x = uninsured,y = totaldeath))

```

```{r}
#plot smoker with death
smokerp <- tablenew1 %>% 
  select(Smoker,Population_Size,Total_Deaths,CHSI_State_Name) %>% 
  mutate(smoker_p = Smoker/100*Population_Size)

table.state3 <- smokerp %>% group_by(CHSI_State_Name) %>% summarise(Total_Deaths = sum(Total_Deaths),smoker_p=sum(smoker_p))

ggplot(data = tablenew1) + 
  geom_point(mapping = aes(x = Smoker, y = Total_Deaths))+ geom_smooth(mapping = aes(x = Smoker, y = Total_Deaths)) + ggtitle( "Mark： Smoker vs Death")

ggplot(table.state3,aes(x = smoker_p, y = Total_Deaths)) + 
  geom_point(aes(color = smoker_p)) + 
  geom_smooth() + ggtitle( "Mark： Smoker vs Death")

```


```{r}
#plot diabetes with death
diabetesp <- tablenew1 %>% 
  select(Diabetes,Population_Size,Total_Deaths,CHSI_State_Name) %>% 
  mutate(diabetes_p = Diabetes/100*Population_Size)

table.state4 <- diabetesp %>% group_by(CHSI_State_Name) %>% summarise(Total_Deaths = sum(Total_Deaths),diabetes_p=sum(diabetes_p))

ggplot(data = tablenew1,mapping = aes(x = Diabetes, y = Total_Deaths)) + 
  geom_point()+ geom_smooth(method = "gam")+
  ggtitle("Grace: Diabetes VS total_death")

ggplot(table.state4,aes(x = diabetes_p, y = Total_Deaths)) + 
  geom_point(aes(color = diabetes_p)) + 
  geom_smooth() +ggtitle("Grace: Diabetes VS total_death")
```


 
  In conclusion, uninsurence, diabetes, smoking and obesity have positive relation with total death counts. Among all the variables, smoking is the most obvious factor that affect the death rate. More smoking causes higher death counts.
  
  



  



```{r}
 data <- data.frame(murder = USArrests$Murder, state = tolower(rownames(USArrests)))
 map <- map_data("state")
 l <- ggplot(data, aes(fill = murder))
 l + geom_map(aes(map_id = state), map = map) +
 expand_limits(x = map$long, y = map$lat)
# map_id, alpha, color, fill, linetype, size)

```
  
  